Reducing retrieval error in MAIAC AOD with the `Earth_obs_cleaning` machine-learning workflow: Results from CONUS

### 2023-08-26 writing/CONUS_AOD.qmd
Remove SHAP table, add captions, remove hyphens in distances.

The study region was the conterminous United States (CONUS) and all match-ups of MODIS data were based on the 1-km grid cells in the Global sinusoidal Coordinate Reference System. The study period included all retrievals from `r min(Wf$dates)` through `r max(Wf$dates)`. The `Earth_obs_cleaning` workflow automatically downloaded all intersecting tiles of MCD19A2 Collection 6.1[@lyapustinalexei2018] based on a NASA-supplied MODIS tile boundary file. While the MCD19A2 product includes data from both Aqua and Terra MODIS instruments, we focus our analyses on the longer running Terra.

@@

MAIAC AOD retrievals from all overpasses in the overlying 1-km grid cell were matched to the closest AERONET observation in time (no more than 8 minutes). Because the purpose of the `Earth_obs_cleaning` workflow is to reduce retrieval error, there was no restriction based on quality codes of input MAIAC data. The targeted estimand in all analyses was the difference of the two AODs (*diff~AOD~*): MAIAC AOD minus AERONET observation, our empirical estimate of retrieval error.

@@

-   the mean AOD calculated from a square window with an 11-km side-length

-   the proportion missing AOD from a square window with an-11 km side-length

@@

After collocation there were `r scales::comma(nrow(dt))` matched AOD retrievals from `r uniqueN(dt$site)` AERONET sites in CONUS. The quartiles of the absolute difference from the labeled overpass time (one per 1200-km wide tile) and the time of the matched AERONET observation were `r tqs = round(quantile(dt[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["25%"]` s, `r tqs["50%"]` s, and `r tqs["75%"]` s. While the distribution of AOD is skewed, our estimate of retrieval error and the dependent variable for our predictive model, *diff~AOD~*, was fairly symmetric and Gaussian (see @fig-dv-density).

@@

In decreasing order of variable importance (from `r the.sat` collocated cross-validation predictions)

```{r}
# what do the SHAPs tell us?
tar_load(cv)
shaps <- cv$"shap_score"
# sort by importance
df_shap <- data.frame(meanabsshap = round(sort(sapply(abs(shaps), mean), decreasing = TRUE), 4))
flextable(df_shap)
# df_shap$variable <- row.names(df_shap)
# ft_shap <- flextable(df_shap[, c(2, 1)])
# ft_shap <- set_header_labels(ft_shap, list(variable = "Variable", meanabsshap = "Mean Absolute SHAP"))
# ft_shap
```

Summary visualization of SHAP contributions

```{r shap_summary}
IG(tar_read(shap.summary.plot.path))
```

```{r shap_features}
IG(tar_read(shap.features.plot.path))
```

@@

### Median MSE day

This map shows a tile on `r tar_read(median.mse.map.data)$date`.

```{r pred.map.median.mse}
IG(tar_read(pred.map.median.mse.path))
```

### Baltimore

A day with long-range transported smoke from Canadian wildfires.

```{r pred.map.baltimore}
IG(tar_read(pred.map.baltimore.path))
```

@@

## Performance by year

```{r}
flextable(pretty.table.numbers(tar_read(cv.summary)[-1]))
```

## Performance by number of overpasses in collocation that site-day

```{r}
dt[, overpasses := .N, by = .(site, lubridate::as_date(time.ground))]
by.overpasses = dt[, eval(performance.j), keyby = overpasses]
flextable(pretty.table.numbers(by.overpasses))
```

>>>>

The study region was the conterminous United States (CONUS) and all match-ups of MODIS data were based on the 1 km grid cells in the Global sinusoidal Coordinate Reference System. The study period included all retrievals from `r min(Wf$dates)` through `r max(Wf$dates)`. The `Earth_obs_cleaning` workflow automatically downloaded all intersecting tiles of MCD19A2 Collection 6.1[@lyapustinalexei2018] based on a NASA-supplied MODIS tile boundary file. While the MCD19A2 product includes data from both Aqua and Terra MODIS instruments, we focus our analyses on the longer running Terra.

@@

MAIAC AOD retrievals from all overpasses in the overlying 1 km grid cell were matched to the closest AERONET observation in time (no more than 8 minutes). Because the purpose of the `Earth_obs_cleaning` workflow is to reduce retrieval error, there was no restriction based on quality codes of input MAIAC data. The targeted estimand in all analyses was the difference of the two AODs (*diff~AOD~*): MAIAC AOD minus AERONET observation, our empirical estimate of retrieval error.

@@

-   the mean AOD calculated from a square window with an 11 km side-length

-   the proportion missing AOD from a square window with an 11 km side-length

@@

After collocation there were `r scales::comma(nrow(dt))` matched AOD retrievals from `r uniqueN(dt$site)` AERONET sites in CONUS. The quartiles of the absolute difference from the labeled overpass time (one per 1200 km wide tile) and the time of the matched AERONET observation were `r tqs = round(quantile(dt[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["25%"]` s, `r tqs["50%"]` s, and `r tqs["75%"]` s. While the distribution of AOD is skewed, our estimate of retrieval error and the dependent variable for our predictive model, *diff~AOD~*, was fairly symmetric and Gaussian (see @fig-dv-density).

@@

```{r shap_summary}
#| label: fig-shap-summary
#| fig-cap: Summary of SHAP contributions. Predictors ordered by decending mean absolute SHAP (printed along left).
IG(tar_read(shap.summary.plot.path))
```

```{r shap_features}
#| label: fig-shap-timeseries
#| fig-cap: Time series of SHAPs for Column Water Vapor (CWV). SHAPs for observations with missing values for CWV are shown in grey.
IG(tar_read(shap.features.plot.path))
```

@@

### Mapping a typical day

@fig-map-median-mse shows a tile on `r tar_read(median.mse.map.data)$date`.

```{r pred.map.median.mse}
#| label: fig-map-median-mse
#| fig-cap: Maps of typical MAIAC and Corrected AOD. Day selected as having the median improvement in MSE at AERONET sites.
IG(tar_read(pred.map.median.mse.path))
```

### Mapping a documented high AOD event

```{r pred.map.baltimore}
#| label: fig-map-baltimore-smoke
#| fig-cap: Maps of MAIAC and Corrected AOD over the Mid-Atlantic US during a documented exceedence event with downmixing of long-range transported smoke from Canadian wildfires.
IG(tar_read(pred.map.baltimore.path))
```

@@

```{r}
#| label: tbl-cv-by-year
#| tbl-cap: Cross-validation performance by year.
flextable(pretty.table.numbers(tar_read(cv.summary)[-1]))
```

```{r}
#| label: tbl-cv-overpasses
#| tbl-cap: Cross-validation performance by the number of overpasses in collocation for that site-day.
dt[, overpasses := .N, by = .(site, lubridate::as_date(time.ground))]
by.overpasses = dt[, eval(performance.j), keyby = overpasses]
flextable(pretty.table.numbers(by.overpasses))
```
